Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-11 WSSSPE1_report/WSSSPE1_summary_report.tex
minor edits to new text

Hanwell et al.~\cite{Hanwell_WSSSPE} reflect on the 15-year history of open
source software development at Kitware.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions. This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables Kitware to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

Foster et al.~\cite{Foster_WSSSPE} discussed the approach of developing sustainability models
around Software-as-a-Service (SaaS) platforms, with the target example being
that of GlobusOnline.  The authors build a case that both grant-based and
volunteer-based development fall short in sustaining software, resulting in
software that is disproportionately difficult to use compared to its
functionality, which they note directly impacts the overall scientific
productivity of its users.  In contrast, by charging a subscription fee for
hosted, centrally-managed software (similar to offerings by Dropbox, EverNote,
GMail), the authors propose to manage the funding cycle and enable a greater
focus on the aspects of software that directly impact individuals, rather than
funders.  Globus has deployed such a service, for which they have attempted to
develop a sustainable economic model that reduces institutional obstacles to
funding and subscription.  However, they do identify that cultural obstacles
do still remain, and they note a particular difference in culture between NSF-
and NIH-funded researchers.

>>>>

Hanwell et al.~\cite{Hanwell_WSSSPE} reflect on the 15-year history of open
source software development at \toolname{Kitware}.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions. This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables \toolname{Kitware} to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

Foster et al.~\cite{Foster_WSSSPE} discussed the approach of developing sustainability models
around Software-as-a-Service (SaaS) platforms, with the target example being
that of \toolname{GlobusOnline}.  The authors build a case that both grant-based and
volunteer-based development fall short in sustaining software, resulting in
software that is disproportionately difficult to use compared to its
functionality, which they note directly impacts the overall scientific
productivity of its users.  In contrast, by charging a subscription fee for
hosted, centrally-managed software (similar to offerings by \toolname{Dropbox}, \toolname{EverNote},
\toolname{GMail}), the authors propose to manage the funding cycle and enable a greater
focus on the aspects of software that directly impact individuals, rather than
funders.  \toolname{Globus} has deployed such a service, for which they have attempted to
develop a sustainable economic model that reduces institutional obstacles to
funding and subscription.  However, they do identify that cultural obstacles
do still remain, and they note a particular difference in culture between NSF-
and NIH-funded researchers.

While many papers briefly discussed licensing issues, Elster~\cite{Elster_WSSSPE}
directly discussed the impact of different licenses on obtaining industrial
funding for scientific software projects.  In particular, she describes
experiences with researchers unwilling to utilize GPL (copyleft) code, as it
adds restrictions to reuse that some industries find unacceptable for future
commercialization.  (This was discussed by Hanwell et al.~\cite{Hanwell_WSSSPE} as well.)
Further obstacles to sustainability identified by the author include patenting
of software, a lack of open access, and even more broadly, a lack of open
source codes even in open access journals.  Finally, a lack of awareness on the
part of scientific software developers of commodity libraries for common tasks
reduces their ability to reuse code.
